---
title: "Esgap_Wolrd"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ESGAP WOLRD

This document contains and allows you to follow the evolution of the code used to process the data in order to produce the Esgap at global level.

## Step 1: Load the necessary packages
```{r warning=FALSE,message=FALSE}
# load packages
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

################################################################################################################################################################

## step 2: process the various data into the desired form

In this step, we format the data tables into a standard form containing the columns countries, variables, years 2015 to 2020.

```{r warning=FALSE,message=FALSE}
Countries<-read.csv("Countries.csv",sep=";")
```
The first column corresponds to the names that will be used for the SESI calculation, and each subsequent column corresponds to the names of the countries in the x database.This table will enable us to standardize the country names for each indicator. This will avoid any loss of information when calculating the SESI index.

!! remember to transform the "," in some databases into " !!

### First topic: Renewable resources

#### **Forest resources** 
Indicator no.	01
Function : Source functions
Principle	: Renew renewable resources
Topic :	Biomass
```{r warning=FALSE,message=FALSE}
Tree<-read.csv("tree_coverage.csv",sep=";")%>%
    select(Country,Variable,c(4:9))

#These lines are used to standardize country names.
correspondances <- match(Tree$Country, Countries$Tree)
Tree$Country<-Countries$Name[correspondances]
Tree <- Tree[complete.cases(Tree$Country), ]
#

Biome<-read.csv("Biome_foret.csv",sep=";")%>%
  group_by(Country,layer_id)%>%
  filter(layer_id=="Boreal" | layer_id=="Temperate" | layer_id=="Tropical")%>%
  summarize(area=sum(area))%>%
  ungroup(Country,layer_id)

correspondances <- match(Biome$Country, Countries$Biome)
Biome$Country<-Countries$Name[correspondances]
Biome <- Biome[complete.cases(Biome$Country), ]

#1
Biome<-full_join(Biome,Biome%>%
  group_by(Country)%>%
  summarize(area_tot=sum(area)))%>%
  mutate(area_per=(area/area_tot))

Tree<-full_join(Tree,Biome)%>%
  mutate(area_2015=as.numeric(X2015)*area_per,
         area_2016=as.numeric(X2016)*area_per,
         area_2017=as.numeric(X2017)*area_per,
         area_2018=as.numeric(X2018)*area_per,
         area_2019=as.numeric(X2019)*area_per,
         area_2020=as.numeric(X2020)*area_per)
#


#2
Tree<-Tree%>%
  group_by(Country,layer_id)%>%
  summarize(X2015=area_2015/area*100,
            X2016=area_2016/area*100,
            X2017=area_2017/area*100,
            X2018=area_2018/area*100,
            X2019=area_2019/area*100,
            X2020=area_2020/area*100)%>% 
  mutate(X2015=ifelse(layer_id=="Boreal",
                ifelse(X2015<60,0,
                    ifelse(X2015>85,100,
                           (X2015-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2015<30,0,
                    ifelse(X2015>50,100,
                           (X2015-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2015<60,0,
                    ifelse(X2015>85,100,
                           (X2015-85)/(60-85))),NA))))%>%
  
   mutate(X2016=ifelse(layer_id=="Boreal",
                ifelse(X2016<60,0,
                    ifelse(X2016>85,100,
                           (X2016-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2016<30,0,
                    ifelse(X2016>50,100,
                           (X2016-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2016<60,0,
                    ifelse(X2016>85,100,
                           (X2016-85)/(60-85))),NA))))%>%
  
   mutate(X2017=ifelse(layer_id=="Boreal",
                ifelse(X2017<60,0,
                    ifelse(X2017>85,100,
                           (X2017-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2017<30,0,
                    ifelse(X2017>50,100,
                           (X2017-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2017<60,0,
                    ifelse(X2017>85,100,
                           (X2017-85)/(60-85))),NA))))%>%
  
   mutate(X2018=ifelse(layer_id=="Boreal",
                ifelse(X2018<60,0,
                    ifelse(X2018>85,100,
                           (X2018-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2018<30,0,
                    ifelse(X2018>50,100,
                           (X2018-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2018<60,0,
                    ifelse(X2018>85,100,
                           (X2018-85)/(60-85))),NA))))%>%
  
   mutate(X2019=ifelse(layer_id=="Boreal",
                ifelse(X2019<60,0,
                    ifelse(X2019>85,100,
                           (X2019-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2019<30,0,
                    ifelse(X2019>50,100,
                           (X2019-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2019<60,0,
                    ifelse(X2019>85,100,
                           (X2019-85)/(60-85))),NA))))%>%
  
   mutate(X2020=ifelse(layer_id=="Boreal",
                ifelse(X2020<60,0,
                    ifelse(X2020>85,100,
                           (X2020-85)/(60-85))),
               ifelse(layer_id=="Temperate",
                ifelse(X2020<30,0,
                    ifelse(X2020>50,100,
                           (X2020-50)/(30-50))),        
               ifelse(layer_id=="Tropical",
                ifelse(X2020<60,0,
                    ifelse(X2020>85,100,
                           (X2020-85)/(60-85))),NA))))%>%
  ungroup()
  
  Tree<-Tree%>%
    group_by(Country)%>%
    summarise(X2015=mean(X2015,na.rm=TRUE),
              X2016=mean(X2016,na.rm=TRUE),
              X2017=mean(X2017,na.rm=TRUE),
              X2018=mean(X2018,na.rm=TRUE),
              X2019=mean(X2019,na.rm=TRUE),
              X2020=mean(X2020,na.rm=TRUE))
#
Tree<-as.data.frame(apply(Tree, 2, as.character))

```

!! Not Working !!

The "Tree" table corresponds to FAO data on forest cover for each country for the years 2015 to 2020, and the "Biome" table corresponds to Global Ecosystem Typology (GET) data on functional forest cover. Both data sets are expressed in square kilometers (km2).

(1)Our objective is to obtain the percentage of forest cover in relation to the cover of functional forest biomes. Initially, we had data on total forest cover in km2 without biome differentiation for the years 2015-2020, and on functional biome cover in km2. To achieve our objective, we need to determine the forest cover of the different functional biomes for each year. To do this, we will apply the percentage of each biome from the GET table to the annual forest cover data, giving us the forest cover of each specific biome by year.

(2)With these new values, we can now compare the difference in forest cover between the functional biomes and the forest cover for the selected years.

#### **Fish resources** 
Indicator no.	02
Function :	Source functions
sub-function :	Renew renewable resources
Topic :	Biomass
```{r warning=FALSE,message=FALSE}
fishsust<-read.csv("Fish_msy_region.csv", sep = ";")%>%
  group_by(Country)%>%
  #1
  summarize(X2017=mean(X2017, na.rm=TRUE),
            X2015=mean(X2015,na.rm=TRUE))%>%
  ungroup()%>%
  #
  
  #2
  mutate(X2016=X2015,
         X2018=X2017,
         X2019=X2017,
         X2020=X2017)
  #
correspondances <- match(fishsust$Country, Countries$Fishstut)
fishsust$Country<-Countries$Name[correspondances]
fishsust <- fishsust[complete.cases(fishsust$Country), ]
fishsust<-as.data.frame(apply(fishsust, 2, as.character))
```
The table has been compiled by hand from data on the proportion of the stock not overexploited (indicator 14.4.1) by major FAO fishing region.

(1) To obtain data by country, we averaged all the regions in which a country could fish.

(2) For years with no value, we use the nearest previous year's value if possible, or the nearest year's value if this is not possible.

#### **Freshwater resources**
Indicator no.	03
Function : Source functions
sub-function : Renew renewable resources
Topic :	Freshwater
```{r warning=FALSE,message=FALSE}
waterstress<-read.csv("Waterstress.csv",sep=";")
waterstress<-waterstress %>%
  select(Country,c(11:16))%>%
  group_by(Country)%>%
  #1
  summarize(X2015=mean(X2015,na.rm=TRUE),
            X2016=mean(X2016,na.rm=TRUE),
            X2017=mean(X2017,na.rm=TRUE),
            X2018=mean(X2018,na.rm=TRUE),
            X2019=mean(X2019,na.rm=TRUE),
            X2020=mean(X2020,na.rm=TRUE))%>%
  ungroup
  #

correspondances <- match(waterstress$Country, Countries$Waterstress)
waterstress$Country<-Countries$Name[correspondances]
waterstress <- waterstress[complete.cases(waterstress$Country), ]
waterstress<-as.data.frame(apply(waterstress, 2, as.character))
```
The table was retrieved from the FAO website (indicator 6.4.2), which corresponds to the proportion of water used in relation to the quantity of water available in the country.

(1) Some countries have changed their calculation method from one year to the next, and therefore appear on several lines. Generally speaking, the values change little or not at all, so we have decided to group everything together on a single line, averaging the two methods if two values exist for the country and the year. If only one value exists, it is taken as it is.

#### **Groundwater resources**
Indicator no.	04
Function : Source functions
sub-function : Renew renewable resources
Topic	: Freshwater
```{r warning=FALSE,message=FALSE}

#\Groundwater produced internally
prodwat<-read.csv("produced_groundwater.csv",sep = ";")%>%
  group_by(Country)%>%
  summarize(X2015=sum(X2015,na.rm = TRUE),X2016=sum(X2016,na.rm= TRUE),X2017=sum(X2017,na.rm= TRUE),X2018=sum(X2018,na.rm= TRUE),X2019=sum(X2019,na.rm= TRUE),X2020=sum(X2020,na.rm= TRUE))


correspondances <- match(prodwat$Country, Countries$produced_GW)
prodwat$Country<-Countries$Name[correspondances]
prodwat <- prodwat[complete.cases(prodwat$Country), ]

#\Groundwater: accounted inflow
inflowwat<-read.csv("inflow_wat.csv",sep=";")%>% 
  group_by(Country)%>%
  summarize(Y2015=sum(X2015,na.rm = TRUE),Y2016=sum(X2016,na.rm= TRUE),Y2017=sum(X2017,na.rm= TRUE),Y2018=sum(X2018,na.rm= TRUE),Y2019=sum(X2019,na.rm= TRUE),Y2020=sum(X2020,na.rm= TRUE))


correspondances <- match(inflowwat$Country, Countries$inflow_GW)
inflowwat$Country<-Countries$Name[correspondances]
inflowwat <- inflowwat[complete.cases(inflowwat$Country), ]

#1
prodwat<-full_join(prodwat,inflowwat)%>%  
  group_by(Country)%>%
  summarize(X2015=Y2015+X2015,
            X2016=Y2016+X2016,
            X2017=Y2017+X2017,
            X2018=Y2018+X2018,
            X2019=Y2019+X2019,
            X2020=Y2020+X2020)%>%
  ungroup()
#

#\Fresh groundwater withdrawal
freshwat<-read.csv("Fresh_groundwater.csv",sep = ";")%>% 
  group_by(Country)%>%
  #2
  summarize(Y2015=sum(X2015,na.rm = TRUE),
            Y2016=sum(X2016,na.rm= TRUE),
            Y2017=sum(X2017,na.rm= TRUE),
            Y2018=sum(X2018,na.rm= TRUE),
            Y2019=sum(X2019,na.rm= TRUE),
            Y2020=sum(X2020,na.rm= TRUE))
  #

correspondances <- match(prodwat$Country, Countries$inflow_GW)
prodwat$Country<-Countries$Name[correspondances]
prodwat <- prodwat[complete.cases(prodwat$Country), ]

groundwat<-full_join(prodwat,freshwat)%>%
  group_by(Country)%>%
  #3
  summarize(X2015=Y2015/X2015*100,
            X2016=Y2016/X2016*100,
            X2017=Y2017/X2017*100,
            X2018=Y2018/X2018*100,
            X2019=Y2019/X2019*100,
            X2020=Y2020/X2020*100)%>%
  ungroup()
  #

groundwat<-as.data.frame(apply(groundwat, 2, as.character))
```
three data sets have been extracted from Aquastat : -Fresh groundwater withdrawal: groundwater use in 10^9m3/year
                                                    -Groundwater accounted inflow: quantity of groundwater entering the country in 10^9m3/year
                                                    -Groundwater produced internally: quantity of groundwater renewed internally in 10^9m3/year.
                                                    
(1) addition of the two groundwater production tables

(2) grouping of values (reason similar to surface water)

(3) ratio of groundwater used to groundwater produced.

#### **Soil resources**
Indicator no.	05
Function	: Source functions
sub-function	: Use non-renewables prudently*
Topic	: Soil
```{r warning=FALSE,message=FALSE}
soil<-read.csv("GlobalErosion_2001_2012.csv",sep=";")%>%
  select(2:4)%>%
  #1
  reframe(Country=Name,
         X2015=MEAN_2012,
         X2016=MEAN_2012,
         X2017=MEAN_2012,
         X2018=MEAN_2012,
         X2019=MEAN_2012,
         X2020=MEAN_2012,)
  #

correspondances <- match(soil$Country, Countries$Erosion)
soil$Country<-Countries$Name[correspondances]
soil <- soil[complete.cases(soil$Country), ]
soil<-as.data.frame(apply(soil, 2, as.character))
```
The table was taken from the ESDAC website and corresponds to the amount of soil erosion for the years 2001 and 2012.

(1) We are interested in the years 2015 to 2020, but no more recent data were found. We therefore use 2012 as the value for the years of interest.

################################################################################################################################################################

### Second topic : Sink function
#### **Climate change**
Indicator no.	06
Function : Sink functions
sub-function : Prevent global warming and the depletion of the ozone layer
Topic	: Earth System 
```{r warning=FALSE,message=FALSE}
#\greenhouse gas (GHG) emissions by country
GHG<-read.csv("GHG_emission.csv",sep=";")

correspondances <- match(GHG$Country, Countries$GHG)
GHG$Country<-Countries$Name[correspondances]
GHG <- GHG[complete.cases(GHG$Country), ]

#\population by country
Pop<-read.csv("population.csv",sep=";")%>%
  select(1,5:10)

colnames(Pop)[1:7]=c("Country","Y2015","Y2016","Y2017","Y2018","Y2019","Y2020")

correspondances <- match(Pop$Country, Countries$population)
Pop$Country<-Countries$Name[correspondances]
Pop <- Pop[complete.cases(Pop$Country), ]

#1
GHG<-left_join(GHG,Pop)%>%
  group_by(Country)%>%
    summarize(X2015=mean(as.numeric(X2015*10^6)/as.numeric(Y2015)), 
              X2016=mean(as.numeric(X2016*10^6)/as.numeric(Y2016)),
              X2017=mean(as.numeric(X2017*10^6)/as.numeric(Y2017)), 
              X2018=mean(as.numeric(X2018*10^6)/as.numeric(Y2018)),  
              X2019=mean(as.numeric(X2019*10^6)/as.numeric(Y2019)),
              X2020=mean(as.numeric(X2020*10^6)/as.numeric(Y2020)))%>%
  ungroup()
#

GHG<-as.data.frame(apply(GHG, 2, as.character))
```
The data was extracted from ESDAC for greenhouse gases and from the World Bank for population per country. The greenhouse gas table is in millions of tonnes per year.

(1) We transform data into tCO_2 per capita per year. We therefore divide GHG emissions by the population of each country.

#### **Depletion of the ozone layer**
Indicator no.	07
Function : Sink functions
sub-function : Prevent global warming and the depletion of the ozone layer
Topic :	Earth System
```{r warning=FALSE,message=FALSE}
HCFC<-read.csv("HCFC_tot.csv",sep=";")

HCFC<-HCFC[-1,]%>%
  select(1,2,10:15)%>%
  #1
  mutate(X=ifelse(is.na(X) | X==0,0,0.005*as.numeric(X)))
  #

colnames(HCFC)=c("Country","X2007","X2015","X2016","X2017","X2018","X2019","X2020")
correspondances <- match(HCFC$Country, Countries$HCFC)
HCFC$Country<-Countries$Name[correspondances]
HCFC <- HCFC[complete.cases(HCFC$Country), ]
HCFC<-as.data.frame(apply(HCFC, 2, as.character))

#\Tableau agreement limit(2)
control<-read.csv("Control_limHCFC.csv",sep=";")
correspondances <- match(control$Country, Countries$Lim_HCFC)
control$Country<-Countries$Name[correspondances]
control <- control[complete.cases(control$Country), ]
control<-as.data.frame(apply(control, 2, as.character))
#
```
Data extracted from the Ozone Secretariat. The table is in HCFC production in ODP tonnes.

(1)Unlike the other indicators, the ideal and critical states are defined at country level and not at global level. We calculate the country's ideal standard, which corresponds to 5% of 2007 consumption.

(2) This data set corresponds to the critical state of each country.

#### **Critical levels in terrestrial ecosystems: ozone**
Indicator no.	08
Function : Sink functions
sub-function : Respect critical levels and loads for ecosystems
Topic	: Terrestrial ecosystems
```{r warning=FALSE,message=FALSE}
#not yet entered
```

#### **Critical loads in terrestrial ecosystems: heavy metals**
Indicator no.	09
Function : Sink functions
sub-function : Respect critical levels and loads for ecosystems
Topic : Terrestrial ecosystems
```{r warning=FALSE,message=FALSE}
#not yet entered
```

#### **Critical loads in terrestrial ecosystems: eutrophication**
Indicator no.	10
Function : Sink functions
sub-function : Respect critical levels and loads for ecosystems
Topic : Terrestrial ecosystems
```{r warning=FALSE,message=FALSE}
#not yet entered
```

#### **Critical loads in terrestrial ecosystems: acidification**
Indicator no.	11
Function : Sink functions
sub-function : Respect critical levels and loads for ecosystems
Topic : Terrestrial ecosystems
```{r warning=FALSE,message=FALSE}
#not yet entered
```

#### **chemical status of surface water bodies**
Indicator no.	12
Function : Sink functions
sub-function : Respect critical loads for ecosystems
Topic : Freshwater ecosystems
```{r warning=FALSE,message=FALSE}
SW_chem<-read.csv("SW_quality.csv",sep=";")

SW_chem<-SW_chem%>%
  select(7:9,22)%>%
  pivot_wider(names_from = Year, values_from = Value)
colnames(SW_chem)=c("Country","Variable","X2017","X2020")
SW_chem<-SW_chem%>%
  group_by(Country)%>%
  summarize(#1
            X2017=mean(X2017,na.rm=TRUE),
            X2020=mean(X2020,na.rm=TRUE))%>%
            #
  #2
  mutate(X2015=X2017,
         X2016=X2017,
         X2018=X2017,
         X2019=X2017)
  #
correspondances <- match(SW_chem$Country, Countries$SW_chem)
SW_chem$Country<-Countries$Name[correspondances]
SW_chem <- SW_chem[complete.cases(SW_chem$Country), ]
SW_chem<-as.data.frame(apply(SW_chem, 2, as.character))
```
Surface water data for the years 2017 and 2020 have been extracted from GEMS (Global Environment Monitoring Systems) and correspond to indicator 6.3.2 of the SDGs.

(1)Rivers and water bodies are differentiated in each country's data. Thus, we average the two types of water surface to obtain a single value per country and per year.

(2)For years with no value, we use the value of the nearest previous year if possible, or the value of the nearest year if this is not possible.

#### **Chemical status of groundwater bodies**
Indicator no.	13
Function : Sink functions
sub-function : Respect critical loads for ecosystems
Topic :	Freshwater ecosystems
```{r warning=FALSE,message=FALSE}
GW_chem<-read.csv("GW_quality.csv",sep=",")

GW_chem<-GW_chem%>%
  select(7:9)
GW_chem<-GW_chem%>%
  distinct()%>%
  pivot_wider(names_from = Year, values_from = Value)

colnames(GW_chem)=c("Country","X2017","X2020","SUP")
#1
GW_chem<-GW_chem%>%
  select(-contains("SUP"))%>%
  mutate(X2015=X2017,
         X2016=X2017,
         X2018=X2017,
         X2019=X2017)
#
correspondances <- match(GW_chem$Country, Countries$GW_chem)
GW_chem$Country<-Countries$Name[correspondances]
GW_chem <- GW_chem[complete.cases(GW_chem$Country), ]
GW_chem<-as.data.frame(apply(GW_chem, 2, as.character))
```
Groundwater data for the years 2017 and 2020 have been extracted from GEMS (Global Environment Monitoring Systems) and correspond to indicator 6.3.2 of the SDGs.

(1)For years with no value, we use the nearest previous year's value if possible, or the nearest year's value if this is not possible.

#### **Chemical status of coastal water **
Indicator no.	14
Function : Sink functions
sub-function : Respect critical levels and loads for ecosystems
Topic : Marine ecosystems
```{r warning=FALSE,message=FALSE}
#pas fait
```

##################################################################################################################################################################
### Troisième topic: Life support

#### **Ecosystem health of terrestrial ecosystems**
Indicator no.	15
Function : Life support functions
sub-function :	Maintain biodiversity and ecosystem health
Topic :	Terrestrial ecosystems
```{r warning=FALSE,message=FALSE}
ter_conv<-read.csv("BII_value.csv",sep=";")%>%
  #1
  mutate(X2015=X2015*100,
         X2016=X2016*100,
         X2017=X2017*100,
         X2018=X2018*100,
         X2019=X2019*100,
         X2020=X2020*100)
  #

correspondances <- match(ter_conv$Country, Countries$BII)
ter_conv$Country<-Countries$Name[correspondances]
ter_conv <- ter_conv[complete.cases(ter_conv$Country), ]
ter_conv<-as.data.frame(apply(ter_conv, 2, as.character))
```
The data table has been compiled by hand, using values from the biodiversity risk assessment tool for AFD countries, produced by Biotope in 2023. The source data for this tool comes from the Muséum d'Histoire Naturelle.

(1)To obtain a percentage

#### **Ecosystem health of freshwater ecosystems**
Indicator no.	16
Function : Life support
sub-function : Maintain biodiversity and ecosystem health
Topic	: Freshwater ecosystems
```{r warning=FALSE,message=FALSE}
#pas fait
```


#### **Ecosystem health of coastal ecosystems**
Indicator no.	17
Function : Life support functions
sub-function : Maintain biodiversity and ecosystem health
Topic :	Marine ecosystems
```{r warning=FALSE,message=FALSE}
Cobio<-read.csv("coast_bio.csv",sep = ",")%>%
  pivot_wider(names_from = scenario, values_from = value)%>%
  filter(long_goal=="Biodiversity" & dimension=="score")%>%
  select(5,9:14)
colnames(Cobio)=c("Country","X2015","X2016","X2017","X2018","X2019","X2020")

correspondances <- match(Cobio$Country, Countries$Coast_bio)
Cobio$Country<-Countries$Name[correspondances]
Cobio<- Cobio[complete.cases(Cobio$Country), ]
Cobio<-as.data.frame(apply(Cobio, 2, as.character))
```
The data were extracted from the IHO website.

(1)the database contains a huge amount of information. We filter only the scores obtained for the IHO Biodiversity indicator. 

##################################################################################################################################################################
### Quatrième topic : Human health and welfare

#### **Outdoor air pollution**
Indicator no.	18
Function : Human health and welfare functions
Principle :	Respect standards for human health
Topic : Human health
```{r warning=FALSE,message=FALSE}
PM2<-read.csv("PM2.5.csv",sep=";")

PM2<-PM2%>%select(Location,Period,FactValueNumeric)%>%
  pivot_wider(names_from = Period, values_from = FactValueNumeric)
colnames(PM2)=c("Country","X2019","X2018","X2017","X2016","X2015")

#1
PM2<-PM2%>%
  mutate(X2020=X2019)
#
correspondances <- match(PM2$Country, Countries$PM2.5)
PM2$Country<-Countries$Name[correspondances]
PM2<- PM2[complete.cases(PM2$Country), ]
PM2<-as.data.frame(apply(PM2, 2, as.character))
```
The dataset comes from the WHO and corresponds to SDG 11.6.2. The low proxy corresponds to 5μg/m3 and the high standard corresponds to 25μg/m3.

(1)For 2020, a year with no value, we use the value for 2019.

#### **Drink water contaminations**
Indicator no.	19
Function : Human health and welfare functions
sub-function : Respect standards for human health
Topic : Human health
```{r warning=FALSE,message=FALSE}
Wash<-read.csv("washdash.csv", sep= ";")
Wash<-Wash%>%group_by(Country,Year,Coverage)%>%
  #1
  filter(Safely.managed.element=="Free from contamination")%>%
  #
  select(Country,Year,Coverage)%>%
  ungroup()%>%
  pivot_wider(names_from = Year, values_from = Coverage)%>%
  select(1,7:12)

colnames(Wash)=c("Country","X2015","X2016","X2017","X2018","X2019","X2020")
correspondances <- match(Wash$Country, Countries$Drinking)
Wash$Country<-Countries$Name[correspondances]
Wash<- Wash[complete.cases(Wash$Country), ]
Wash<-as.data.frame(apply(Wash, 2, as.character))
```
The dataset corresponds to indicator 6.1.1 retrieved from washdata (Analyse by safely managed criteria).

(1)We keep only the data corresponding to "Free from contamination" water, which corresponds to our indicator "free from contamination by Escherichia coli (E.coli) in any 100 ml sample".

#### **Natural and mixed world heritage sites**
Indicator no.	20
Function : Human health and welfare functions
sub-function : Conserve landscape and amenity
Topic : Other welfare
```{r warning=FALSE,message=FALSE}
Worldher<-read.csv("Worldheri.csv",sep = ";")

Worldher<-Worldher%>%
  group_by(Country)%>%
  #1
  summarise(X2014=mean((1-X2014/4)*100,na.rm=TRUE),
            X2017=mean((1-X2017/4)*100,na.rm=TRUE),
            X2020=mean((1-X2020/4)*100,na.rm=TRUE))%>% 
  #
  #2
  rename("X2015"=X2014)%>%
  mutate(X2016=X2015,
         X2018=X2017,
         X2019=X2017)%>%
  #
  ungroup

correspondances <- match(Worldher$Country, Countries$Heritage)
Worldher$Country<-Countries$Name[correspondances]
Worldher<- Worldher[complete.cases(Worldher$Country), ]
Worldher<-as.data.frame(apply(Worldher, 2, as.character))
```
The dataset had to be made by hand and is taken from the Heritage Outlook documents 2014, 2017 and 2020. The data corresponds to the condition of the site and varies between 1 and 4. 1 being good condition and 4 poor.

(1)We first pass all our values as a percentage and then average the site values for each country to obtain a single value. 1 corresponds to good condition and 4 to bad, so we need to take the inverse of the calculated result.

(2)Since we don't have values for all years, we take the values for the closest preceding year.

## step 3: obtain a table grouping all indicators by country and globally (before standardization)

###function to group indicators by country
```{r warning=FALSE,message=FALSE}
country_table<-function(name){

  #\Source Function
  #1
  Tr<-Tree%>%
    filter(Country==name)%>%
    mutate(Function="Source",
           Indicator="state of forest reccources",
           ideal_state="100",
           unit="%")
  #
  
  Fs<-fishsust %>%
    filter(Country==name)%>%
    mutate(Function="Source",
           Indicator="Proportion of fish stocks not overexploited",
           ideal_state="100",
           unit="% Sus fish stock")
     
  Ws<-waterstress%>%
    filter(Country==name)%>% 
    mutate(Function="Source",
           Indicator="Surfacewater Stress",
           ideal_state="25",
           unit="% of stress") 
      
  Gwt<-groundwat%>%
    filter(Country==name)%>%
    mutate(Function="Source",
           Indicator="Groundwater stress",
           ideal_state="25",
           unit="%")
  
  Soil<-soil%>%
    filter(Country==name)%>%
    mutate(Function="Source",
           Indicator="area with tolerable erosion",
           ideal_state="2",
           unit="Tonnes/hectare/year")
   
  #\Sink function
  ghg<-GHG%>%
    filter(Country==name)%>%
    mutate(Function="Sink",
           Indicator="CO2 emissions [tonnes per capita]",
           ideal_state="2.88",
           unit="tonnes per year per capita")
   
  hcfc<-HCFC%>%
    filter(Country==name)%>%
    mutate(Function="Sink",
           Indicator="Consumption of HCFC in ODP per capita",
           unit="0.5% compare to 2007 in CO2-eq tonnes")%>%
           rename("ideal_state"=X2007)
  
  Schem<-SW_chem%>%
    filter(Country==name)%>%
    mutate(Function="Sink",
           Indicator="Proportion of surface water in good chemical status",
           ideal_state="100",
           unit="%")
  
  Gchem<-GW_chem%>%
    filter(Country==name)%>%
    mutate(Function="Sink",
           Indicator="Proportion of ground water in good chemical status",
           ideal_state="100",
           unit="%")
  
  #\Life support
  BII<-ter_conv%>%
    filter(Country==name)%>%
    mutate(Function="Life support",
           Indicator="Terrestrial habitats in favourable conservation status",
           ideal_state="100",
           unit="%")
  
  
  Cobi<-Cobio%>%
    filter(Country==name)%>%
    mutate(Function="Life support",
           Indicator="Aquatic habitats in favourable conservation status",
           ideal_state="100",
           unit="%")
  
  #\Human health and welfare function
  pm<-PM2%>%
    filter(Country==name)%>%
    mutate(Function="Human health and welfare",
           Indicator="Concentrations of PM2.5",
           ideal_state="5",
           unit="μg/m3")
  
  Drink<-Wash%>%
    filter(Country==name)%>%
    mutate(Function="Human health and welfare",
           Indicator="Drinking water contamination",
           ideal_state="100",
           unit="% without E.coli")
  
  Wher<-Worldher%>%
    filter(Country==name)%>%
    mutate(Function="Human health and welfare",
           Indicator="state of World heritage sites",
           ideal_state="100",
           unit="% Good state (2015<-value of 2014)")

  Capnt=bind_rows(Tr,Fs,Ws,Gwt,Soil,ghg,hcfc,Schem,Gchem,BII,Cobi,pm,Drink,Wher)
  #2  
  write.csv(Capnt,paste0("Country/",name,".csv"))
  #
}
```
This function groups together all the indicators for which we have data for a country. If we don't have data for the indicator for the country, it won't appear in the table. 

(1)This section retrieves information from the Tree table corresponding to the Forest resources indicator for the country named 'name'. In addition, 4 informative columns are added, indicating the function of the indicator, 

(2)The table is saved on the computer in csv format in the file "/Country".

##Table example
```{r warning=FALSE,message=FALSE}
country_table("Brazil")
test_indic<-read.csv("Country/Brazil.csv")
test_indic
```
We have chosen Brazil as our country because all the indicators are complete.

###function for a table of all countries
This section allows you to create a table containing all data by country and by indicator.
```{r warning=FALSE,message=FALSE}
all_indic<-function(Countries){
  #1
  for (i in 1:nrow(Countries)){
    n<-Countries[i,1]
    country_table(n)
  #
  }
  #2
  all_countries<-read.csv(paste0("Country/",Countries[1,1],".csv"))
  for (i in 2:nrow(Countries)){
    n<-Countries[i,1]
    y<-read.csv(paste0("Country/",n,".csv"))
    all_countries<-rbind(all_countries,y)
  #
  }
  #3
  write.csv(all_countries,paste0("Country/","all_countries",".csv"))
  #
}
```
(1)Repeats the precedent function to create a table for each country, based on the standardized names, and saves it in the Country folder.

(2)This part reopens each table for each country created previously and links them together.

(3)Saves the table in the Country folder, as for tables for one country only.

##Table of all countries
```{r warning=FALSE,message=FALSE}
all_indic(Countries)
all<-read.csv("Country/all_countries.csv")
```

##step 4: normalize data and calculate SESI

###function to normalize and calculate SESi by country
```{r warning=FALSE,message=FALSE}
#1
normhab<-function(value,gpmin,gpmax){5+(gpmax-5)*((value-gpmin)/(gpmax-gpmin))}
#

normall<-function(name)
{
  Pays<-read.csv(paste0("Country/",name,".csv"))
  table<-read.csv(paste0("Country/",name,".csv"))
  n<-1
      
  ###NORMALISER###
  
  ##Source Function
  #forest
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Forest reccources"){
   for (i in 4:9){
         x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }
  
  #fish :
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Proportion of fish stocks not overexploited"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }
  
  #water stress:
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Surface water Stress"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-ifelse(!is.na(x) & x<=25,100,ifelse(!is.na(x) & x>=100,5,abs(normhab(x,25,100))))
    }
    n<-n+1
  }
  #
  
  #Groundwater
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Groundwater stress"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-ifelse(!is.na(x) & x>=100,5,ifelse(!is.na(x) & x<=25,100,abs(normhab(x,25,100))))
    }
    n<-n+1
  }
  
  #Erosion

  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="area with tolerable erosion"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-ifelse(x<2,100,
                    ifelse(x>10,5,normhab(((x-10)/(2-10))*100,0,100)))
    }
    n<-n+1
  }
  
  ##Sink function
  #GHG
    if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="CO2 emissions [tonnes per capita]"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-ifelse(x<2,100,ifelse(x>6,5,normhab((1-(x-2)/(6-2))*100,0,100)))
      }
    n<-n+1
    }
  
  #HCFC
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Consumption of HCFC in ODP per capita"){
    control2<-control%>%
      filter(Country==name)
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE) 
      y<-as.numeric(control2[i-2])
      prox<-ifelse(as.numeric(Pays$proxy[n])>y,5,as.numeric(Pays$proxy[n]))
      x<-((x-y)/(prox-y))*100
      table[n,i]<-ifelse(x<0.1,5,ifelse(x>100,100,normhab(x,0,100)))
    
      }
    n<-n+1
  }
  #
  
  #surface water in good chemical status :
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Proportion of surface water bodies in good chemical status"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }
   
  #groundwater in good chemical status :
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Proportion of ground water bodies in good chemical status"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }
  
  ##life support function
  #Biodiv Terrestrial
      if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Terrestrial habitats in favourable conservation status"){
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-ifelse(x>=90,100,
                         ifelse(x<=30,5,
                                normhab(x,30,90)))
    }
    n<-n+1
      }
  ##life support function
  #Biodiv Aquatic
    if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Aquatic habitats in favourable conservation status"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }
  
  ##human Health function
  #Pm2.5
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Concentrations of PM2.5"){ 
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      x<-(x-5)/(35-5)*100
      table[n,i]<-ifelse(x<0,5,normhab(x,0,100))
    }
    n<-n+1
  }
  
  #drinking water
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="Drinking water"){
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      table[n,i]<-normhab(x,0,100)
    }
    n<-n+1
  }

  #World Heritage:
  if (!is.na(Pays$Variable[n]) & Pays$Variable[n]=="World heritage"){
    for (i in 4:9){
      x<-as.numeric(Pays[n,i],na.rm=TRUE)
      x<-normhab(x,0,100)
      table[n,i]<-x
    }
    n<-n+1
  }

  table<-table%>%
    select(2:10)
  #
  SESI<-NULL
  for (i in 3:8){
    SESI<-c(SESI,prod(as.numeric(table[c(1:nrow(Pays)),i]),na.rm=TRUE)^(1/sum(!is.na(as.numeric(table[c(1:nrow(Pays)),i])))))
    
  }
  SESI<-c(name,"SESI",SESI,n)
  #
  table<-rbind(table,SESI)
  write.csv(table,paste0("Country_normSESI/",name,".csv"))
}
```
The table is saved in the Country_normSESI folder.

(1) normalization function for the various indicators. Due to the use of a geometric mean to calculate the SESI, it is impossible to have values equal to 0. The lowest value for normalization is therefore 5, as described by Usubiaga-Liaño and Ekin. 

(Water stress)Values above the ideal state are directly normalized to 100% and values below the critical state are directly normalized to 5%.

(HCFC)As explained above, ideal and critical states are defined at country level for this indicator. It is therefore necessary to recover them.

(2)We calculate the SESI, which corresponds to the geometric mean of all indicators for a given year.

##Table example
```{r warning=FALSE,message=FALSE}
normall("Brazil")
test_norm<-read.csv("Country_normSESI/Brazil.csv")
test_norm
```


##Table of all normalize countries
Table containing all the standardized data and SESI calculations for which we have found data for each country. A table is saved for each country in doc Country_normSESI
```{r warning=FALSE,message=FALSE}
all_norm<-function(Countries){
  #to obtain a table for each country
  for (i in 1:nrow(Countries)){
    n<-Countries[i,1]
    normall(n)
    print(n)
  }
  #to obtain a table of all countries
  all_normcountries<-read.csv(paste0("Country_normSESI/",Countries[1,1],".csv"))
  for (i in 2:nrow(Countries)){
    n<-Countries[i,1]
    y<-read.csv(paste0("Country_normSESI/",n,".csv"))
    all_normcountries<-full_join(all_normcountries,y)
  }
  write.csv(all_normcountries,paste0("Country_normSESI/","all_normcountries",".csv"))
}
all_norm(Countries)

all_norm<-read.csv("Country_normSESI/all_normcountries.csv")
```
This part is a function similar to the one above for obtaining the table grouping all data by country and by indicator. The table is saved in the Country_normSESI folder.



